---
title: "Welcome to tidyterra"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Welcome to tidyterra}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warning = FALSE,
  message = FALSE,
  fig.align = "center"
)
```

## Welcome to {tidyterra}

**tidyterra** is a package that add common methods from the
[tidyverse](https://www.tidyverse.org/) for SpatRaster and SpatVectors objects
created with the [{terra}](https://CRAN.R-project.org/package=terra) package. It
also adds specific`geom_spat*()` for plotting these kind of objects with
[{ggplot2}](https://ggplot2.tidyverse.org/).

### Why {tidyterra}?

Spat\* objects are not like regular data frames. They are a different type of
objects, implemented via the [S4 object system](http://adv-r.had.co.nz/S4.html),
and have their own syntax and computation methods, implemented on the {terra}
package.

By implementing tidyverse methods for these objects, and more specifically
{dplyr} and {tidyr} methods, a use**R** can now work more easily with Spat\*,
just like (s)he would do with tabular data.

**Note that** in terms of performance, {terra} is much more optimized for
working for this kind of objects, so it is **recommended** also to learn a bit
of {terra} syntax. Each function of {tidyterra} refers (when possible) to the
corresponding equivalent on {terra}.

## A note for advanced {terra} users

As previously mentioned, {tidyterra} is not optimized in terms of performance.
Specially when working with `filter()` and `mutate()` methods, it can be slow.

As a rule of thumb, {tidyterra} can handle objects with less than 10.000.000
slots of information(i.e., `terra::ncell(a_rast) * terra::nlyr(a_rast) < 10e6`).

## Get started with {tidyterra}

First thing you'll notice when loading {tidyterra} is that a set of packages
would be also loaded. You can disable the message by setting
`Sys.setenv(tidyterra.quiet = TRUE)`.

```{r setup, message=TRUE}
# Disable message with:
# Sys.setenv(tidyterra.quiet = TRUE)
library(tidyterra)
```

Currently, the following methods are available:

| tidyverse method      | SpatVector | SpatRaster      |
|-----------------------|------------|-----------------|
| `tibble::as_tibble()` | ✔️         | ✔️              |
| `dplyr::filter()`     | ✔️         | ✔️              |
| `dplyr::mutate()`     | ✔️         | ✔️              |
| `dplyr::pull()`       | ✔️         | ✔️              |
| `dplyr::relocate()`   | ✔️         | ✔️              |
| `dplyr::rename()`     | ✔️         | ✔️              |
| `dplyr::select()`     | ✔️         | ✔️              |
| `dplyr::slice()`      | ✔️         | ✔️              |
| `dplyr::transmute()`  | ✔️         | ✔️              |
| `tidyr::drop_na()`    | ✔️         | ✔️ (questioned) |
| `tidyr::replace_na()` | ✔️         | ✔️              |

Let's see some of them in action:

```{r}
library(terra)
f <- system.file("extdata/cyl_temp.tif", package = "tidyterra")

temp <- rast(f)

temp

mod <- temp %>%
  select(-1) %>%
  mutate(newcol = tavg_06 - tavg_05) %>%
  relocate(newcol, .before = 1) %>%
  replace_na(list(newcol = 3)) %>%
  rename(difference = newcol)

mod
plot(mod)
```

On the previous example, we had:

-   Eliminated the first layer of the raster `tavg_04`.

-   Created a new layer `newcol` as the difference of the layers `tavg_05` and
    `tavg_06.`

-   Relocated `newcol`as the first layer of the SpatRaster

-   Replaced the `NA` cells on `newcol` with `3`.

-   Renamed `newcol` to difference.

In all the process, the essential properties of the SpatRaster (number of cells,
columns and rows, extent, resolution and coordinate reference system) have not
been modified. Other methods as `filter()`, `slice()` or `drop_na()` can modify
these properties, as they would do when applied to a data frame (number of rows
would be modified on that case).

## Plotting with {ggplot2}

### SpatRasters

{tidyterra} provides several `geom_*` for SpatRasters. When the SpatRaster has
the CRS informed (i.e. `terra::crs(a_rast) != ""`), the geom uses
`ggplot2::coord_sf()`, and may be also reprojected for adjusting the coordinates
to other spatial layers:

```{r terraplots}

library(ggplot2)

# A faceted SpatRaster

ggplot() +
  geom_spatraster(data = temp) +
  facet_wrap(~lyr) +
  scale_fill_gradientn(
    colours = hcl.colors(10, "RdBu", rev = TRUE),
    na.value = "white"
  )


# Contour lines for a specific layer

f_volcano <- system.file("extdata/volcano2.tif", package = "tidyterra")
volcano2 <- rast(f_volcano)

ggplot() +
  geom_spatraster(data = volcano2) +
  geom_spatraster_contour(data = volcano2, breaks = seq(80, 200, 5)) +
  scale_fill_topo_c() +
  coord_sf(expand = FALSE) +
  labs(fill = "elevation")


# Contour filled

ggplot() +
  geom_spatraster_contour_filled(data = volcano2) +
  scale_fill_terrain_d(direction = -1) +
  labs(fill = "elevation")
```

With {tidyterra} you can also plot RGB SpatRasters to add imagery to your plots:

```{r rgb}

# Read a vector

f_v <- system.file("extdata/cyl.gpkg", package = "tidyterra")
v <- vect(f_v)

# Read a tile
f_rgb <- system.file("extdata/cyl_tile.tif", package = "tidyterra")

r_rgb <- rast(f_rgb)

rgb_plot <- ggplot() +
  geom_spatraster_rgb(data = r_rgb) +
  geom_spatvector(data = v, fill = NA, size = 1)

rgb_plot
# Change CRS automatically

rgb_plot +
  coord_sf(crs = 3035)
```

### SpatVectors

{tidyterra} allows you to plot SpatVectors with {ggplot2} using the
`geom_spatvector()` functions:

```{r}
lux <- system.file("ex/lux.shp", package = "terra")

spatvector <- terra::vect(lux)

ggplot() +
  geom_spatvector(data = spatvector, aes(fill = POP)) +
  geom_spatvector_text(
    data = spatvector, aes(label = NAME_2),
    color = "grey90"
  ) +
  scale_fill_binned(labels = scales::number_format()) +
  coord_sf(crs = 3857)
```

The underlying implementation is to take advantage of the conversion
`terra::vect()/sf::st_as_sf()` (see [**About SpatVectors**](#about)) and use
`ggplot2::geom_sf()` as an endpoint for creating the layer.

### About SpatVectors {#about}

SpatVector objects are vector data. This means that they are a set of individual
points with geographic information (i.e. location of restaurants), that can be
also grouped to form lines (i.e. a river) and, when these lines forms a closed
polygon, a spatial polygon (i.e. a country).

{terra} can handle vector files on the S4 system. There is other alternative,
the {sf} package, that represents the same information on a tabular way. You can
convert easily between the two packages like this:

```{r terra_sf}
lux <- system.file("ex/lux.shp", package = "terra")

spatvector <- terra::vect(lux)
spatvector

terra::plot(spatvector, main = "SpatVector", axes = TRUE)


# To sf
sfobj <- sf::st_as_sf(spatvector)
head(sfobj, 3)

plot(sfobj$geometry, main = "sf", axes = TRUE)

# Back to terra

spatvector2 <- terra::vect(sfobj)

spatvector2
```

On that sense, {sf} already has its own implementation of {tidyverse} methods.
Since converting SpatVector to {sf} it is straightforward with no loss of
information, {tidyterra} is specially focused on SpatRasters. See an example of
how to work with {terra} + {sf} + {dplyr}:

```{r full_spatvector}
library(dplyr)
library(terra)
library(sf)

spat_summ <- spatvector2 %>%
  # to sf
  st_as_sf() %>%
  # dplyr
  group_by(ID_1) %>%
  summarise(TOTPOP = sum(POP)) %>%
  # back to terra
  vect()

spat_summ

terra::plot(spat_summ, "TOTPOP")
```
