---
title: "tidyterra FAQs"
description: Frequently Asked Questions on the use of tidyterra.
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warning = FALSE,
  message = FALSE,
  dpi = 300,
  cache = FALSE,
  dev = "ragg_png",
  tidy = "styler",
  out.width = "100%"
)
```

This is a compendium of [questions
arisen](https://github.com/dieghernan/tidyterra/discussions) on the use of the
**tidyterra** package and the potential solutions to it (mostly related with the
use of **terra** and **ggplot2** at this stage). You can ask for help or search
previous questions in the following links.

You can also ask in [Stack Overflow](https://stackoverflow.com/) using the tag
[`tidyterra`](https://stackoverflow.com/questions/tagged/tidyterra).

```{r issues, echo=FALSE, results='asis'}
link <- unlist(packageDescription("tidyterra")["BugReports"])
cat("\n- Issues: ", link)
dis <- gsub("issues$", "discussions", link)
cat("\n- Discussions: ", dis)
cat("\n")
```

## `NA` values are shown in gray color {#nas-remove}

This is the default behavior produced by the **ggplot2** package. **tidyterra**
color scales (i.e., `scale_fill_whitebox_c()`, etc.), has by default the
parameter `na.value` set to `"transparent"`, that prevents `NA` values to be
filled[^1].

[^1]: `na.value = NA` could be used as well for the same purpose in most of the
    cases, however when the proportion of non-`NA`s is small it can produce
    undesired results, see
    [#120](https://github.com/dieghernan/tidyterra/issues/120).

```{r remove-nas}
library(terra)
library(tidyterra)
library(ggplot2)

# Get a raster data, out beloved volcano on hi-res
volcanotemp <- "volcano2hires.tif"

# Download example file
volcanourl <- paste0(
  "https://github.com/dieghernan/tidyterra/blob/main/",
  "data-raw/volcano2hires.tif?raw=true"
)

if (!file.exists(volcanotemp)) {
  download.file(volcanourl, volcanotemp, mode = "wb")
}
r <- volcanotemp %>%
  rast() %>%
  filter(elevation > 80 & elevation < 180)

# Default
def <- ggplot() +
  geom_spatraster(data = r)

def +
  labs(
    title = "Default on ggplot2",
    subtitle = "NA values in grey"
  )

# Modify with scales
def +
  scale_fill_continuous(na.value = "transparent") +
  labs(
    title = "Default colors on ggplot2",
    subtitle = "But NAs are not plotted"
  )

# Use a different scale provided by ggplot2
def +
  scale_fill_viridis_c(na.value = "orange") +
  labs(
    title = "Use any fill_* scale of ggplot2",
    subtitle = "Note that na.value = 'orange'"
  )
```

## Labeling contours {#label-contour}

Thanks to `fortify.SpatRaster()` you can use your `SpatRaster` straight away
with the **metR** package. Use the parameter(s) `bins/binwidth/breaks` to align
both labels and lines:

```{r text-contour}
library(terra)
library(tidyterra)
library(ggplot2)
library(metR)

volcanotemp <- "volcano2hires.tif"

# Download example file
volcanourl <- paste0(
  "https://github.com/dieghernan/tidyterra/blob/main/",
  "data-raw/volcano2hires.tif?raw=true"
)

if (!file.exists(volcanotemp)) {
  download.file(volcanourl, volcanotemp, mode = "wb")
}


r <- rast(volcanotemp)

ggplot(r) +
  geom_spatraster_contour(data = r) +
  geom_text_contour(
    aes(x, y, z = elevation),
    check_overlap = TRUE,
    stroke = 0.2,
    stroke.colour = "white"
  ) +
  labs(
    title = "Labelling contours",
    width = 2,
    x = "", y = ""
  )

# Modify number or bins

ggplot(r) +
  geom_spatraster_contour(
    data = r,
    binwidth = 25
  ) +
  geom_text_contour(
    aes(x, y, z = elevation),
    binwidth = 25,
    check_overlap = TRUE,
    stroke = 0.2, stroke.colour = "white"
  ) +
  labs(
    title = "Labelling contours",
    subtitle = "Aligning breaks",
    width = 2,
    x = "", y = ""
  )
```

## Using a different color scale {#use-scale}

Since **tidyterra** leverages on **ggplot2**, please refer to **ggplot2** use of
scales:

```{r greys}
library(terra)
library(tidyterra)
library(ggplot2)

volcanotemp <- "volcano2hires.tif"

# Download example file
volcanourl <- paste0(
  "https://github.com/dieghernan/tidyterra/blob/main/",
  "data-raw/volcano2hires.tif?raw=true"
)

if (!file.exists(volcanotemp)) {
  download.file(volcanourl, volcanotemp, mode = "wb")
}


r <- rast(volcanotemp)

# Hillshade with grey colors
slope <- terrain(r, "slope", unit = "radians")
aspect <- terrain(r, "aspect", unit = "radians")
hill <- shade(slope, aspect, 10, 340)

ggplot() +
  geom_spatraster(data = hill, show.legend = FALSE) +
  # Note the scale, grey colours
  scale_fill_gradientn(
    colours = grey(0:100 / 100),
    na.value = "transparent"
  ) +
  labs(title = "A hillshade plot with grey colors")
```

## Can I change the default palette of my maps?

Yes, use `options("ggplot2.continuous.fill")` to modify the default colors on
your session.

```{r default}
library(terra)
library(tidyterra)
library(ggplot2)

volcanotemp <- "volcano2hires.tif"

# Download example file
volcanourl <- paste0(
  "https://github.com/dieghernan/tidyterra/blob/main/",
  "data-raw/volcano2hires.tif?raw=true"
)

if (!file.exists(volcanotemp)) {
  download.file(volcanourl, volcanotemp, mode = "wb")
}


r <- rast(volcanotemp)

p <- ggplot() +
  geom_spatraster(data = r)


# Set options
tmp <- getOption("ggplot2.continuous.fill") # store current setting
options(ggplot2.continuous.fill = scale_fill_terrain_c)

p

# restore previous setting
options(ggplot2.continuous.fill = tmp)


p
```

## My map tiles are blurry {#blurry-tiles}

This is probably related with the tile itself rather than the package. Most base
tiles are provided in **EPSG:3857**, so check first if your tile has this CRS
and not a different one. Not having **EPSG:3857** may be an indication that the
tile has been reprojected, implied some sort of sampling that causes the
blurriness on your data. Also, modify the parameter `maxcell` to avoid
resampling and force the **ggplot2** map to be on **EPSG:3857** with
`ggplot2::coord_sf(crs = 3857)`:

```{r blurry-tile}
library(terra)
library(tidyterra)
library(ggplot2)
library(sf)
library(maptiles)

# Get a tile from a point on sf format

p <- st_point(c(-97.09, 31.53)) %>%
  st_sfc(crs = 4326) %>%
  st_buffer(750)

tile1 <- get_tiles(p, provider = "OpenStreetMap", zoom = 14, cachedir = ".")

ggplot() +
  geom_spatraster_rgb(data = tile1) +
  geom_sf(data = p, fill = NA) +
  labs(title = "This is a bit blurry...")

st_crs(tile1)$epsg

# The tile was in EPSG 4326

# get tile in 3857
p2 <- st_transform(p, 3857)


tile2 <- get_tiles(p2, provider = "OpenStreetMap", zoom = 14, cachedir = ".")

st_crs(tile2)$epsg

# Now the tile is EPSG:3857

ggplot() +
  geom_spatraster_rgb(data = tile2, maxcell = Inf) +
  geom_sf(data = p, fill = NA) +
  # Force crs to be 3857
  coord_sf(crs = 3857) +
  labs(
    title = "See the difference?",
    subtitle = "Init crs=3857 and maxcell modified"
  )
```

## Avoid degrees labeling on axis {#axis-degrees}

Again, this is the **ggplot2** default, but can be modified with
`ggplot2::coord_sf(datum)` argument:

```{r modify-datum}
library(terra)
library(tidyterra)
library(ggplot2)
library(sf)

volcanotemp <- "volcano2hires.tif"

# Download example file
volcanourl <- paste0(
  "https://github.com/dieghernan/tidyterra/blob/main/",
  "data-raw/volcano2hires.tif?raw=true"
)

if (!file.exists(volcanotemp)) {
  download.file(volcanourl, volcanotemp, mode = "wb")
}


r <- rast(volcanotemp)

ggplot() +
  geom_spatraster(data = r) +
  labs(
    title = "Axis auto-converted to lon/lat",
    subtitle = paste("But SpatRaster is EPSG:", st_crs(r)$epsg)
  )


# Use datum

ggplot() +
  geom_spatraster(data = r) +
  coord_sf(datum = pull_crs(r)) +
  labs(
    title = "Axis on the units of the SpatRaster",
    subtitle = paste("EPSG:", st_crs(r)$epsg)
  )
```

## Modifying the number of breaks on axis {#axis-breaks}

The best option is to pass your custom breaks to `ggplot2::scale_x_continous()`
or `ggplot2::scale_y_continous()`. You will need to provide the breaks in
lon/lat even if your data is projected. See also
[ggplot2/issues/4622](https://github.com/tidyverse/ggplot2/issues/4622):

```{r breaks}
library(terra)
library(tidyterra)
library(ggplot2)
library(sf)

volcanotemp <- "volcano2hires.tif"

# Download example file
volcanourl <- paste0(
  "https://github.com/dieghernan/tidyterra/blob/main/",
  "data-raw/volcano2hires.tif?raw=true"
)

if (!file.exists(volcanotemp)) {
  download.file(volcanourl, volcanotemp, mode = "wb")
}

r <- rast(volcanotemp)

ggplot() +
  geom_spatraster(data = r) +
  labs(title = "Default axis breaks")

# Modify y breaks with extent projected in EPSG:4326

# Get extent

ext <- r %>%
  project("EPSG:4326", mask = TRUE) %>%
  ext() %>%
  as.vector()


ggplot() +
  geom_spatraster(data = r) +
  scale_y_continuous(
    expand = expansion(mult = 0.05),
    breaks = scales::breaks_pretty(n = 3)(ext[c("ymin", "ymax")])
  ) +
  labs(title = "Three breaks on y")
```

## Plotting a `SpatRaster` with color tables

**tidyterra** has several ways to handle these `SpatRaster` objects:

```{r coltab}
library(terra)
library(tidyterra)
library(ggplot2)

# Get a SpatRaster with coltab
r_coltab <- rast(system.file("extdata/cyl_era.tif", package = "tidyterra"))

has.colors(r_coltab)

r_coltab

# Native handling by terra packages
plot(r_coltab)


# A. autoplot

autoplot(r_coltab, maxcell = Inf) + ggtitle("autoplot method")

# B. geom_spatraster
ggplot() +
  geom_spatraster(data = r_coltab, maxcell = Inf) +
  ggtitle("geom_spatraster method")

# C. Using scale_fill_coltab

g <- ggplot() +
  geom_spatraster(data = r_coltab, use_coltab = FALSE, maxcell = Inf)

g

# But...
g +
  scale_fill_coltab(data = r_coltab) +
  ggtitle("scale_fill_coltab method")

# D. Extract named colors and scale_fill_manual

cols <- get_coltab_pal(r_coltab)

cols

scales::show_col(cols)

# And now

g +
  scale_fill_manual(
    values = cols, na.value = "transparent",
    na.translate = FALSE
  ) +
  ggtitle("scale_fill_manual method")
```

## North arrows and scale bar

**tidyterra** does not provide these graphical objects for **ggplot2** plots.
However, you can use **ggspatial** functions
(`ggspatial::annotation_north_arrow()` and `ggspatial::annotation_scale()`):

```{r northarrow}
library(terra)
library(tidyterra)
library(ggplot2)
library(ggspatial)

volcanotemp <- "volcano2hires.tif"

# Download example file
volcanourl <- paste0(
  "https://github.com/dieghernan/tidyterra/blob/main/",
  "data-raw/volcano2hires.tif?raw=true"
)

if (!file.exists(volcanotemp)) {
  download.file(volcanourl, volcanotemp, mode = "wb")
}

r <- rast(volcanotemp)

autoplot(r) +
  annotation_north_arrow(
    which_north = TRUE,
    pad_x = unit(0.8, "npc"),
    pad_y = unit(0.75, "npc"),
    style = north_arrow_fancy_orienteering()
  ) +
  annotation_scale(
    height = unit(0.015, "npc"),
    width_hint = 0.5,
    pad_x = unit(0.07, "npc"),
    pad_y = unit(0.07, "npc"),
    text_cex = .8
  )
```

## How to overlay a `SpatRaster` over a RGB tile

This is quite straightforward, just use `geom_spatraster_rgb()` and after that
command just create your layer:

```{r overlay_cont}
library(terra)
library(tidyterra)
library(ggplot2)
library(sf)
# Get example data
library(maptiles)
library(geodata)


# Area of interest
aoi <- gadm(country = "CHE", path = ".", level = 0) %>%
  project("EPSG:3857")

# Tile
rgb_tile <- get_tiles(aoi,
  crop = TRUE, provider = "Esri.WorldShadedRelief",
  zoom = 8, project = FALSE, cachedir = "."
)

# Clim (mean prec)
clim <- worldclim_country("CHE", var = "prec", path = ".") %>%
  project(rgb_tile) %>%
  mask(aoi) %>%
  terra::mean()

# Labels
cap_lab <- paste0(
  c(
    "Tiles © Esri - Source: Esri",
    "Data: © Copyright 2020-2022, worldclim.org."
  ),
  collapse = "\n"
)
tit_lab <- "Average precipitation in Switzerland"

ggplot(aoi) +
  geom_spatraster_rgb(data = rgb_tile, alpha = 1) +
  geom_spatraster(data = clim) +
  geom_spatvector(fill = NA) +
  scale_fill_whitebox_c(
    palette = "deep", alpha = 0.5,
    labels = scales::label_number(suffix = " mm.")
  ) +
  coord_sf(expand = FALSE) +
  labs(
    title = tit_lab,
    subtitle = "With continuous overlay",
    fill = "Precipitation",
    caption = cap_lab
  )
```

We can create another variations with binned legends and filled contours (see
`geom_spatraster_contour_filled()`):

```{r overlay_alt}
# Binned
ggplot(aoi) +
  geom_spatraster_rgb(data = rgb_tile, alpha = 1) +
  geom_spatraster(data = clim) +
  geom_spatvector(fill = NA) +
  scale_fill_whitebox_b(
    palette = "deep", alpha = 0.5,
    n.breaks = 4,
    labels = scales::label_number(suffix = " mm.")
  ) +
  coord_sf(expand = FALSE) +
  labs(
    title = tit_lab,
    subtitle = "With overlay: binned legend",
    fill = "Precipitation",
    caption = cap_lab
  )

# Filled contour
ggplot(aoi) +
  geom_spatraster_rgb(data = rgb_tile, alpha = 1) +
  geom_spatraster_contour_filled(data = clim, bins = 4) +
  geom_spatvector(fill = NA) +
  coord_sf(expand = FALSE) +
  scale_fill_whitebox_d(
    palette = "deep", alpha = 0.5,
    guide = guide_legend(reverse = TRUE)
  ) +
  labs(
    title = tit_lab,
    subtitle = "With overlay and filled contour",
    fill = "Precipitation (mm.)",
    caption = cap_lab
  )
```

## Hexagonal grids (and other `geoms`)

By concept the cells of a `SpatRaster` are rectangular, so it is not possible to
create a `SpatRaster` with i.e. hexagonal cells.

But it is possible to create a plot with hexagonal cells thanks to
`fortify.SpatRaster()` and `stat_summary_hex()`. Additional work is needed to
adjust the final plot, specifically it is needed also to use `coord_sf()`:

```{r hex_grid}
library(terra)
library(tidyterra)
library(ggplot2)

volcanotemp <- "volcano2hires.tif"

# Download example file
volcanourl <- paste0(
  "https://github.com/dieghernan/tidyterra/blob/main/",
  "data-raw/volcano2hires.tif?raw=true"
)

if (!file.exists(volcanotemp)) {
  download.file(volcanourl, volcanotemp, mode = "wb")
}

r <- rast(volcanotemp)

# With hex grid
ggplot(r, aes(x, y, z = elevation)) +
  stat_summary_hex(
    fun = mean, color = NA, linewidth = 0,
    # Bins size determines the number of cells displayed
    bins = 30
  ) +
  coord_sf(crs = pull_crs(r)) +
  labs(
    title = "Hexagonal SpatRaster",
    subtitle = "Using fortify (implicit) and stat_summary_hex",
    x = NULL, y = NULL
  )
```

Note that there is no need to make a direct call to `fortify.SpatRaster()`,
since this function is implicitly invoked by **ggplot2** when using
`ggplot(data = a_spatraster)`.

Thanks to this extension mechanism, it is possible to use additional `geoms` and
`stats` provided by **ggplot2**:

```{r alt_points}
# Point plot
ggplot(r, aes(x, y, z = elevation), maxcell = 1000) +
  geom_point(aes(size = elevation, alpha = elevation),
    fill = "darkblue",
    color = "grey50", shape = 21
  ) +
  coord_sf(crs = pull_crs(r)) +
  scale_radius(range = c(1, 5)) +
  scale_alpha(range = c(0.01, 1)) +
  labs(
    title = "SpatRaster as points",
    subtitle = "Using fortify (implicit)",
    x = NULL, y = NULL
  )
```

## Session info

<details>

<summary>Details</summary>

```{r session, echo=FALSE}
if (!require("sessioninfo")) {
  install.packages("sessioninfo")
}
sessioninfo::session_info()
```

</details>
