---
title: "tidyterra FAQs"
description: Frequently Asked Questions on the use of tidyterra.
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warning = FALSE,
  message = FALSE,
  dpi = 300,
  cache = FALSE,
  dev = "ragg_png",
  tidy = "styler",
  out.width = "100%"
)
```

This is a compendium of questions arisen on the use of the **tidyterra** package
and the potential solutions to it (mostly related with the use of **terra** and
**ggplot2** at this stage). You can ask for help or search previous questions in
the following links.

```{r issues, echo=FALSE, results='asis'}
link <- unlist(packageDescription("tidyterra")["BugReports"])
cat("\n- Issues: ", link)
dis <- gsub("issues$", "discussions", link)
cat("\n- Discussions: ", dis)
cat("\n")
```

## NA values are shown in gray color {#nas-remove}

This is the default behavior produced by the **ggplot2** package. **tidyterra**
color scales (i.e., `scale_fill_whitebox_c()`, etc.), has by default the
parameter `na.value` set to `NA`, that prevents NA values to be plotted.

```{r remove-nas}
library(terra)
library(tidyterra)
library(ggplot2)

# Get a raster data, out beloved volcano on hi-res
volcanotemp <- file.path(tempdir(), "volcano2hires.tif")

# Download example file
volcanourl <- paste0(
  "https://github.com/dieghernan/tidyterra/blob/main/",
  "data-raw/volcano2hires.tif?raw=true"
)

if (!file.exists(volcanotemp)) {
  download.file(volcanourl, volcanotemp, mode = "wb")
}
r <- volcanotemp %>%
  rast() %>%
  filter(elevation > 80 & elevation < 180)

# Default
def <- ggplot() +
  geom_spatraster(data = r)

def +
  labs(
    title = "Default on ggplot2",
    subtitle = "NA values in grey"
  )

# Modify with scales
def +
  scale_fill_continuous(na.value = NA) +
  labs(
    title = "Default colors on ggplot2",
    subtitle = "But NAs are not plotted"
  )

# Use a different scale provided by ggplot2
def +
  scale_fill_viridis_c(na.value = "orange") +
  labs(
    title = "Use any fill_* scale of ggplot2",
    subtitle = "Note that na.value = 'orange'"
  )
```

## Labeling contours {#label-contour}

Thanks to `fortify.SpatRaster()` you can use your SpatRaster straight away with
the **metR** package. Use the parameter(s) `bins/binwidth/breaks` to align both
labels and lines:

```{r text-contour}
library(terra)
library(tidyterra)
library(ggplot2)
library(metR)

volcanotemp <- file.path(tempdir(), "volcano2hires.tif")

# Download example file
volcanourl <- paste0(
  "https://github.com/dieghernan/tidyterra/blob/main/",
  "data-raw/volcano2hires.tif?raw=true"
)

if (!file.exists(volcanotemp)) {
  download.file(volcanourl, volcanotemp, mode = "wb")
}


r <- rast(volcanotemp)

ggplot(r) +
  geom_spatraster_contour(data = r) +
  geom_text_contour(
    aes(x, y, z = elevation),
    check_overlap = TRUE,
    stroke = 0.2,
    stroke.colour = "white"
  ) +
  labs(
    title = "Labelling contours",
    width = 2,
    x = "", y = ""
  )

# Modify number or bins

ggplot(r) +
  geom_spatraster_contour(
    data = r,
    binwidth = 25
  ) +
  geom_text_contour(
    aes(x, y, z = elevation),
    binwidth = 25,
    check_overlap = TRUE,
    stroke = 0.2, stroke.colour = "white"
  ) +
  labs(
    title = "Labelling contours",
    subtitle = "Aligning breaks",
    width = 2,
    x = "", y = ""
  )
```

## Using a different color scale {#use-scale}

Since **tidyterra** leverages on **ggplot2**, please refer to **ggplot2** use of
scales:

```{r greys}
library(terra)
library(tidyterra)
library(ggplot2)

volcanotemp <- file.path(tempdir(), "volcano2hires.tif")

# Download example file
volcanourl <- paste0(
  "https://github.com/dieghernan/tidyterra/blob/main/",
  "data-raw/volcano2hires.tif?raw=true"
)

if (!file.exists(volcanotemp)) {
  download.file(volcanourl, volcanotemp, mode = "wb")
}


r <- rast(volcanotemp)

# Hillshade with grey colors
slope <- terrain(r, "slope", unit = "radians")
aspect <- terrain(r, "aspect", unit = "radians")
hill <- shade(slope, aspect, 10, 340)

ggplot() +
  geom_spatraster(data = hill, show.legend = FALSE) +
  # Note the scale, grey colours
  scale_fill_gradientn(colours = grey(0:100 / 100), na.value = NA) +
  labs(title = "A hillshade plot with grey colors")
```

## Can I change the default palette of my maps?

Yes, use `options("ggplot2.continuous.fill")` to modify the default colors on
your session.

```{r default}
library(terra)
library(tidyterra)
library(ggplot2)

volcanotemp <- file.path(tempdir(), "volcano2hires.tif")

# Download example file
volcanourl <- paste0(
  "https://github.com/dieghernan/tidyterra/blob/main/",
  "data-raw/volcano2hires.tif?raw=true"
)

if (!file.exists(volcanotemp)) {
  download.file(volcanourl, volcanotemp, mode = "wb")
}


r <- rast(volcanotemp)

p <- ggplot() +
  geom_spatraster(data = r)


# Set options
tmp <- getOption("ggplot2.continuous.fill") # store current setting
options(ggplot2.continuous.fill = scale_fill_terrain_c)

p

# restore previous setting
options(ggplot2.continuous.fill = tmp)


p
```

## My map tiles are blurry {#blurry-tiles}

This is probably related with the tile itself rather than the package. Most base
tiles are provided in **EPSG:3857**, so check first if your tile has this CRS
and not a different one. Not having **EPSG:3857** may be an indication that the
tile has been reprojected, implied some sort of sampling that causes the
blurriness on your data. Also, modify the parameter `maxcell` to avoid
resampling and force the **ggplot2** map to be on **EPSG:3857** with
`ggplot2::coord_sf(crs = 3857)`:

```{r blurry-tile}
library(terra)
library(tidyterra)
library(ggplot2)
library(sf)
library(maptiles)

# Get a tile from a point on sf format

p <- st_point(c(-97.09, 31.53)) %>%
  st_sfc(crs = 4326) %>%
  st_buffer(750)

tile1 <- get_tiles(p, provider = "Stamen.Terrain", zoom = 14, cachedir = ".")

ggplot() +
  geom_spatraster_rgb(data = tile1) +
  geom_sf(data = p, fill = NA) +
  labs(title = "This is a bit blurry...")

st_crs(tile1)$epsg

# The tile was in EPSG 4326

# get tile in 3857
p2 <- st_transform(p, 3857)


tile2 <- get_tiles(p2, provider = "Stamen.Terrain", zoom = 14, cachedir = ".")

st_crs(tile2)$epsg

# Now the tile is EPSG:3857

ggplot() +
  geom_spatraster_rgb(data = tile2, maxcell = Inf) +
  geom_sf(data = p, fill = NA) +
  # Force crs to be 3857
  coord_sf(crs = 3857) +
  labs(
    title = "See the difference?",
    subtitle = "Init crs=3857 and maxcell modified"
  )
```

## Avoid degrees labeling on axis {#axis-degrees}

Again, this is the **ggplot2** default, but can be modified with
`ggplot2::coord_sf(datum)` argument:

```{r modify-datum}
library(terra)
library(tidyterra)
library(ggplot2)
library(sf)

volcanotemp <- file.path(tempdir(), "volcano2hires.tif")

# Download example file
volcanourl <- paste0(
  "https://github.com/dieghernan/tidyterra/blob/main/",
  "data-raw/volcano2hires.tif?raw=true"
)

if (!file.exists(volcanotemp)) {
  download.file(volcanourl, volcanotemp, mode = "wb")
}


r <- rast(volcanotemp)

ggplot() +
  geom_spatraster(data = r) +
  labs(
    title = "Axis auto-converted to lon/lat",
    subtitle = paste("But SpatRaster is EPSG:", st_crs(r)$epsg)
  )


# Use datum

ggplot() +
  geom_spatraster(data = r) +
  coord_sf(datum = pull_crs(r)) +
  labs(
    title = "Axis on the units of the SpatRaster",
    subtitle = paste("EPSG:", st_crs(r)$epsg)
  )
```

## Modifying the number of breaks on axis {#axis-breaks}

The best option is to pass your custom breaks to `ggplot2::scale_x_continous()`
or `ggplot2::scale_y_continous()`. You will need to provide the breaks in
lon/lat even if your data is projected (¿?):

```{r breaks}
library(terra)
library(tidyterra)
library(ggplot2)
library(sf)

volcanotemp <- file.path(tempdir(), "volcano2hires.tif")

# Download example file
volcanourl <- paste0(
  "https://github.com/dieghernan/tidyterra/blob/main/",
  "data-raw/volcano2hires.tif?raw=true"
)

if (!file.exists(volcanotemp)) {
  download.file(volcanourl, volcanotemp, mode = "wb")
}

r <- rast(volcanotemp)

ggplot() +
  geom_spatraster(data = r) +
  labs(title = "Default axis breaks")

# Modify y breaks

# Get extent

ext <- r %>%
  project("EPSG:4326", mask = TRUE) %>%
  ext() %>%
  as.vector()

br_y <- seq(ext["ymin"], ext["ymax"], length.out = 1000) %>%
  pretty(n = 3) %>%
  round(3) %>%
  unique()

ggplot() +
  geom_spatraster(data = r) +
  scale_y_continuous(
    expand = expansion(mult = 0.05),
    breaks = br_y
  ) +
  labs(title = "Three breaks on y")
```

## Plotting SpatRasters with color tables

**tidyterra** has several ways to handle these SpatRasters:

```{r coltab}
library(terra)
library(tidyterra)
library(ggplot2)

# Get a SpatRaster with coltab
r_coltab <- rast(system.file("raster/nlcd.tif", package = "spDataLarge"))

has.colors(r_coltab)

r_coltab

# Native handling by terra packages
plot(r_coltab)


# A. autoplot

autoplot(r_coltab) + ggtitle("autoplot method")

# B. geom_spatraster
ggplot() +
  geom_spatraster(data = r_coltab) +
  ggtitle("geom_spatraster method")

# C. Using scale_fill_coltab

g <- ggplot() +
  geom_spatraster(data = r_coltab, use_coltab = FALSE)

g

# But...
g +
  scale_fill_coltab(data = r_coltab) +
  ggtitle("scale_fill_coltab method")

# D. Extract named colors and scale_fill_manual

cols <- get_coltab_pal(r_coltab)

cols

scales::show_col(cols)

# And now

g +
  scale_fill_manual(values = cols) +
  ggtitle("scale_fill_manual method")
```
