---
title: "Gradient palettes in tidyterra"
description: A gallery of maps using the gradients provided by tidyterra.
lightbox:
  match: auto
  desc-position: bottom
  effect: fade
format:
  html:
    include-before-body:
      text: |
        <style>
          .glightbox-clean .gslide-desc {
            font-weight: 800;
            font-family: 'Inter Tight';
            text-shadow: 0 0 10px black;
            color: white !important;
            opacity: 0.75;
          }
          .glightbox-mobile .glightbox-container .gslide-desc {
            font-size: 1.25rem;
          }
        </style>
knitr:
  opts_chunk:
    collapse: true
    comment: "#>"
    warning: false
    message: false
    dpi: 300
    tidy: "styler"
    dev: "ragg_png"
    out.width: "100%"
---

This page shows a [Lightbox gallery](https://biati-digital.github.io/glightbox/)
of maps that can be displayed with the gradient fill scales included with
**tidyterra**. A basic code to generate this map is:

```{r}
#| eval: false
library(tidyterra)
library(terra)
library(ggplot2)

r <- rast(system.file("extdata/volcano2.tif", package = "tidyterra"))

ggplot() +
  geom_spatraster(data = r) +
  # Use the palette you like, in this case:
  scale_fill_hypso_c() +
  theme_void()
```

## `scale_fill_terrain_*` and `scale_fill_wiki_*`

Check `scale_fill_terrain_c()` and `scale_fill_wiki_c()` for more info.

```{r}
#| label: terr_wiki
#| echo: false
#| fig-alt: Terrain and wiki palettes in tidyterra
#| layout-ncol: 2
#| lightbox:
#|   group: r-wikiterr
#|   description: Terrain and wiki palettes in tidyterra
library(tidyterra)

r <- terra::rast(system.file("extdata/volcano2.tif", package = "tidyterra"))

library(ggplot2)

baseplot <- ggplot() +
  geom_spatraster(data = r)

ss <- baseplot +
  scale_fill_terrain_c() +
  labs(title = "terrain") +
  theme_void() +
  theme(plot.background = element_rect(fill = "white"))

print(ss)


ss <- baseplot +
  scale_fill_wiki_c() +
  labs(title = "wiki") +
  theme_void() +
  theme(plot.background = element_rect(fill = "white"))


print(ss)
```

## `scale_fill_whitebox_*`

Check `scale_fill_whitebox_c()` for more info.

```{r}
#| label: whitebox
#| echo: false
#| fig-alt: Whitebox palettes in tidyterra
#| layout-ncol: 2
#| lightbox:
#|   group: r-whitebox
#|   description: Whitebox palettes in tidyterra
allpals <- sort(c(
  "atlas",
  "high_relief",
  "arid",
  "soft",
  "muted",
  "purple",
  "viridi",
  "gn_yl",
  "pi_y_g",
  "bl_yl_rd",
  "deep"
))


for (i in allpals) {
  ss <- baseplot +
    scale_fill_whitebox_c(palette = i) +
    labs(title = paste0(i)) +
    theme_void() +
    theme(plot.background = element_rect(fill = "white"))


  print(ss)
}
```

## `scale_fill_hypso_*`

Check `scale_fill_hypso_c()` for more info.

```{r}
#| label: hypso
#| echo: false
#| fig-alt: Hypso palettes in tidyterra
#| layout-ncol: 2
#| lightbox:
#|   group: r-hypso
#|   description: Hypso palettes in tidyterra
allpals <- sort(unique(hypsometric_tints_db$pal))

for (i in allpals) {
  ss <- baseplot +
    scale_fill_hypso_c(palette = i) +
    labs(title = paste0(i)) +
    theme_void() +
    theme(plot.background = element_rect(fill = "white"))


  print(ss)
}
```

## `scale_fill_cross_blended_*`

Check `scale_fill_cross_blended_c()` for more info.

```{r}
#| label: cross
#| echo: false
#| fig-alt: Cross-blended palettes in tidyterra
#| layout-ncol: 2
#| lightbox:
#|   group: r-cross
#|   description: Cross-blended palettes in tidyterra
allpals <- sort(unique(cross_blended_hypsometric_tints_db$pal))

for (i in allpals) {
  ss <- baseplot +
    scale_fill_cross_blended_c(palette = i) +
    labs(title = paste0(i)) +
    theme_void() +
    theme(plot.background = element_rect(fill = "white"))


  print(ss)
}
```

## `scale_fill_grass_*`

Check `scale_fill_grass_c()` for more info. Plots produced using
`use_grass_range = FALSE`.

```{r}
#| label: grass
#| echo: false
#| fig-alt: GRASS palettes in tidyterra
#| layout-ncol: 2
#| lightbox:
#|   group: r-grass
#|   description: GRASS palettes in tidyterra
allpals <- sort(unique(grass_db$pal))

for (i in allpals) {
  ss <- baseplot +
    scale_fill_grass_c(palette = i, use_grass_range = FALSE) +
    labs(title = paste0(i)) +
    theme_void() +
    theme(plot.background = element_rect(fill = "white"))


  print(ss)
}
```

## `scale_fill_princess_*`

Check `scale_fill_princess_c()` for more info.

```{r}
#| label: princess
#| echo: false
#| fig-alt: Princess palettes in tidyterra
#| layout-ncol: 2
#| lightbox:
#|   group: r-princess
#|   description: Princess palettes in tidyterra
allpals <- sort(unique(princess_db$pal))

for (i in allpals) {
  ss <- baseplot +
    scale_fill_princess_c(palette = i) +
    labs(title = paste0(i)) +
    theme_void() +
    theme(plot.background = element_rect(fill = "white"))


  print(ss)
}
```
