---
title: "Gradient palettes in tidyterra"
author: "Diego Hernangómez"
description: >
  A gallery of maps using the gradients provided by tidyterra.
editor_options: 
  markdown: 
    wrap: 80
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warning = FALSE,
  message = FALSE,
  dpi = 300,
  tidy = "styler",
  out.width = "50%",
  fig.show = "hold"
)

library(tidyterra)
```

This page shows a gallery of maps that can be displayed with the gradient fill
scales included with **tidyterra**. A basic code to generate this map is:

```{r, eval=FALSE}

library(tidyterra)
library(terra)
library(ggplot2)

r <- rast(system.file("extdata/volcano2.tif", package = "tidyterra"))

ggplot() +
  geom_spatraster(data = r) +
  # Use the palette you like, in this case:
  scale_fill_hypso_c() +
  theme_void()
```

## `scale_fill_terrain_*` and `scale_fill_wiki_*`

Check `scale_fill_terrain_c()` and `scale_fill_wiki_c()` for more info.

```{r terr_wiki, echo=FALSE}

r <- terra::rast(system.file("extdata/volcano2.tif",
  package = "tidyterra"
))

library(ggplot2)

baseplot <- ggplot() +
  geom_spatraster(data = r)

ss <- baseplot +
  scale_fill_terrain_c() +
  labs(title = "terrain") +
  theme_void()

print(ss)


ss <- baseplot +
  scale_fill_wiki_c() +
  labs(title = "wiki") +
  theme_void()

print(ss)
```

## `scale_fill_whitebox_*`

Check `scale_fill_whitebox_c()` for more info.

```{r whitebox, echo=FALSE}

allpals <- sort(c(
  "atlas", "high_relief", "arid", "soft", "muted", "purple",
  "viridi", "gn_yl", "pi_y_g", "bl_yl_rd", "deep"
))


for (i in allpals) {
  ss <- baseplot +
    scale_fill_whitebox_c(palette = i) +
    labs(title = paste0(i)) +
    theme_void()

  print(ss)
}
```

## `scale_fill_hypso_*`

Check `scale_fill_hypso_c()` for more info.

```{r hypso, echo=FALSE}

allpals <- sort(unique(hypsometric_tints_db$pal))

for (i in allpals) {
  ss <- baseplot +
    scale_fill_hypso_c(palette = i) +
    labs(title = paste0(i)) +
    theme_void()

  print(ss)
}
```

## `scale_fill_cross_blended_*`

Check `scale_cross_blended_hypso_c()` for more info.

```{r cross, echo=FALSE}

allpals <- sort(unique(cross_blended_hypsometric_tints_db$pal))

for (i in allpals) {
  ss <- baseplot +
    scale_fill_cross_blended_c(palette = i) +
    labs(title = paste0(i)) +
    theme_void()

  print(ss)
}
```
