---
title: "Gradient palettes in tidyterra"
description: A gallery of maps using gradient scales provided by **tidyterra**.
lightbox:
  match: auto
  desc-position: bottom
  effect: fade
format:
  html:
    include-before-body:
      text: |
        <style>
          .glightbox-clean .gslide-desc {
            font-weight: 800;
            font-family: 'Inter Tight';
            text-shadow: 0 0 10px black;
            color: white !important;
            opacity: 0.75;
          }
          .glightbox-mobile .glightbox-container .gslide-desc {
            font-size: 1.25rem;
          }
        </style>
knitr:
  opts_chunk:
    collapse: true
    comment: "#>"
    warning: false
    message: false
    dpi: 300
    tidy: "styler"
    dev: "ragg_png"
    out.width: "100%"
---

This page shows a [lightbox gallery](https://biati-digital.github.io/glightbox/)
of maps created with the gradient fill scales included in **tidyterra**. A basic
example for creating similar maps is:

```{r}
#| eval: false
library(tidyterra)
library(terra)
library(ggplot2)

r <- rast(system.file("extdata/volcano2.tif", package = "tidyterra"))

ggplot() +
  geom_spatraster(data = r) +
  # Use the selected palette.
  scale_fill_hypso_c() +
  theme_void()
```

## `scale_fill_terrain_*` and `scale_fill_wiki_*`

See `scale_fill_terrain_c()` and `scale_fill_wiki_c()` for details.

```{r}
#| label: terr_wiki
#| echo: false
#| fig-alt:
#|   - "Raster map of volcanic elevation using the terrain gradient. Fill color encodes elevation across the volcanic terrain."
#|   - "Raster map of volcanic elevation using the wiki gradient. Fill color encodes elevation across the volcanic terrain."
#| layout-ncol: 2
#| lightbox:
#|   group: r-wikiterr
#|   description: Terrain and wiki gradient palettes in tidyterra.
library(tidyterra)

r <- terra::rast(system.file("extdata/volcano2.tif", package = "tidyterra"))

library(ggplot2)

baseplot <- ggplot() +
  geom_spatraster(data = r)

ss <- baseplot +
  scale_fill_terrain_c() +
  labs(title = "terrain") +
  theme_void() +
  theme(plot.background = element_rect(fill = "white"))

print(ss)
ss <- baseplot +
  scale_fill_wiki_c() +
  labs(title = "wiki") +
  theme_void() +
  theme(plot.background = element_rect(fill = "white"))

print(ss)
```

## `scale_fill_whitebox_*`

See `scale_fill_whitebox_c()` for details.

```{r}
#| label: whitebox
#| echo: false
#| fig-alt:
#|   - "Raster map of volcanic elevation using the Whitebox arid palette. Fill color encodes elevation across the volcanic terrain."
#|   - "Raster map of volcanic elevation using the Whitebox atlas palette. Fill color encodes elevation across the volcanic terrain."
#|   - "Raster map of volcanic elevation using the Whitebox bl_yl_rd palette. Fill color encodes elevation across the volcanic terrain."
#|   - "Raster map of volcanic elevation using the Whitebox deep palette. Fill color encodes elevation across the volcanic terrain."
#|   - "Raster map of volcanic elevation using the Whitebox gn_yl palette. Fill color encodes elevation across the volcanic terrain."
#|   - "Raster map of volcanic elevation using the Whitebox high_relief palette. Fill color encodes elevation across the volcanic terrain."
#|   - "Raster map of volcanic elevation using the Whitebox muted palette. Fill color encodes elevation across the volcanic terrain."
#|   - "Raster map of volcanic elevation using the Whitebox pi_y_g palette. Fill color encodes elevation across the volcanic terrain."
#|   - "Raster map of volcanic elevation using the Whitebox purple palette. Fill color encodes elevation across the volcanic terrain."
#|   - "Raster map of volcanic elevation using the Whitebox soft palette. Fill color encodes elevation across the volcanic terrain."
#|   - "Raster map of volcanic elevation using the Whitebox viridi palette. Fill color encodes elevation across the volcanic terrain."
#| layout-ncol: 2
#| lightbox:
#|   group: r-whitebox
#|   description: Whitebox palettes in tidyterra.
allpals <- sort(c(
  "atlas",
  "high_relief",
  "arid",
  "soft",
  "muted",
  "purple",
  "viridi",
  "gn_yl",
  "pi_y_g",
  "bl_yl_rd",
  "deep"
))
for (i in allpals) {
  ss <- baseplot +
    scale_fill_whitebox_c(palette = i) +
    labs(title = paste0(i)) +
    theme_void() +
    theme(plot.background = element_rect(fill = "white"))

  print(ss)
}
```

## `scale_fill_hypso_*`

See `scale_fill_hypso_c()` for details.

```{r}
#| label: hypso
#| echo: false
#| fig-alt:
#|   - "Raster map of volcanic elevation using the hypsometric arctic palette. Fill color encodes elevation across the volcanic terrain."
#|   - "Raster map of volcanic elevation using the hypsometric arctic_bathy palette. Fill color encodes elevation across the volcanic terrain."
#|   - "Raster map of volcanic elevation using the hypsometric arctic_hypso palette. Fill color encodes elevation across the volcanic terrain."
#|   - "Raster map of volcanic elevation using the hypsometric c3t1 palette. Fill color encodes elevation across the volcanic terrain."
#|   - "Raster map of volcanic elevation using the hypsometric colombia palette. Fill color encodes elevation across the volcanic terrain."
#|   - "Raster map of volcanic elevation using the hypsometric colombia_bathy palette. Fill color encodes elevation across the volcanic terrain."
#|   - "Raster map of volcanic elevation using the hypsometric colombia_hypso palette. Fill color encodes elevation across the volcanic terrain."
#|   - "Raster map of volcanic elevation using the hypsometric dem_poster palette. Fill color encodes elevation across the volcanic terrain."
#|   - "Raster map of volcanic elevation using the hypsometric dem_print palette. Fill color encodes elevation across the volcanic terrain."
#|   - "Raster map of volcanic elevation using the hypsometric dem_screen palette. Fill color encodes elevation across the volcanic terrain."
#|   - "Raster map of volcanic elevation using the hypsometric etopo1 palette. Fill color encodes elevation across the volcanic terrain."
#|   - "Raster map of volcanic elevation using the hypsometric etopo1_bathy palette. Fill color encodes elevation across the volcanic terrain."
#|   - "Raster map of volcanic elevation using the hypsometric etopo1_hypso palette. Fill color encodes elevation across the volcanic terrain."
#|   - "Raster map of volcanic elevation using the hypsometric gmt_globe palette. Fill color encodes elevation across the volcanic terrain."
#|   - "Raster map of volcanic elevation using the hypsometric gmt_globe_bathy palette. Fill color encodes elevation across the volcanic terrain."
#|   - "Raster map of volcanic elevation using the hypsometric gmt_globe_hypso palette. Fill color encodes elevation across the volcanic terrain."
#|   - "Raster map of volcanic elevation using the hypsometric meyers palette. Fill color encodes elevation across the volcanic terrain."
#|   - "Raster map of volcanic elevation using the hypsometric meyers_bathy palette. Fill color encodes elevation across the volcanic terrain."
#|   - "Raster map of volcanic elevation using the hypsometric meyers_hypso palette. Fill color encodes elevation across the volcanic terrain."
#|   - "Raster map of volcanic elevation using the hypsometric moon palette. Fill color encodes elevation across the volcanic terrain."
#|   - "Raster map of volcanic elevation using the hypsometric moon_bathy palette. Fill color encodes elevation across the volcanic terrain."
#|   - "Raster map of volcanic elevation using the hypsometric moon_hypso palette. Fill color encodes elevation across the volcanic terrain."
#|   - "Raster map of volcanic elevation using the hypsometric nordisk-familjebok palette. Fill color encodes elevation across the volcanic terrain."
#|   - "Raster map of volcanic elevation using the hypsometric nordisk-familjebok_bathy palette. Fill color encodes elevation across the volcanic terrain."
#|   - "Raster map of volcanic elevation using the hypsometric nordisk-familjebok_hypso palette. Fill color encodes elevation across the volcanic terrain."
#|   - "Raster map of volcanic elevation using the hypsometric pakistan palette. Fill color encodes elevation across the volcanic terrain."
#|   - "Raster map of volcanic elevation using the hypsometric spain palette. Fill color encodes elevation across the volcanic terrain."
#|   - "Raster map of volcanic elevation using the hypsometric usgs-gswa2 palette. Fill color encodes elevation across the volcanic terrain."
#|   - "Raster map of volcanic elevation using the hypsometric utah_1 palette. Fill color encodes elevation across the volcanic terrain."
#|   - "Raster map of volcanic elevation using the hypsometric wiki-2.0 palette. Fill color encodes elevation across the volcanic terrain."
#|   - "Raster map of volcanic elevation using the hypsometric wiki-2.0_bathy palette. Fill color encodes elevation across the volcanic terrain."
#|   - "Raster map of volcanic elevation using the hypsometric wiki-2.0_hypso palette. Fill color encodes elevation across the volcanic terrain."
#|   - "Raster map of volcanic elevation using the hypsometric wiki-schwarzwald-cont palette. Fill color encodes elevation across the volcanic terrain."
#|   - "Raster map of volcanic elevation using the hypsometric xkcd-painbow palette. Fill color encodes elevation across the volcanic terrain."
#| layout-ncol: 2
#| lightbox:
#|   group: r-hypso
#|   description: Hypsometric tint palettes in tidyterra.
allpals <- sort(unique(hypsometric_tints_db$pal))

for (i in allpals) {
  ss <- baseplot +
    scale_fill_hypso_c(palette = i) +
    labs(title = paste0(i)) +
    theme_void() +
    theme(plot.background = element_rect(fill = "white"))

  print(ss)
}
```

## `scale_fill_cross_blended_*`

See `scale_fill_cross_blended_c()` for details.

```{r}
#| label: cross
#| echo: false
#| fig-alt:
#|   - "Raster map of volcanic elevation using the cross-blended arid palette. Fill color encodes elevation across the volcanic terrain."
#|   - "Raster map of volcanic elevation using the cross-blended cold_humid palette. Fill color encodes elevation across the volcanic terrain."
#|   - "Raster map of volcanic elevation using the cross-blended polar palette. Fill color encodes elevation across the volcanic terrain."
#|   - "Raster map of volcanic elevation using the cross-blended warm_humid palette. Fill color encodes elevation across the volcanic terrain."
#| layout-ncol: 2
#| lightbox:
#|   group: r-cross
#|   description: Cross-blended hypsometric tint palettes in tidyterra.
allpals <- sort(unique(cross_blended_hypsometric_tints_db$pal))

for (i in allpals) {
  ss <- baseplot +
    scale_fill_cross_blended_c(palette = i) +
    labs(title = paste0(i)) +
    theme_void() +
    theme(plot.background = element_rect(fill = "white"))

  print(ss)
}
```

## `scale_fill_grass_*`

See `scale_fill_grass_c()` for details. These plots are produced with
`use_grass_range = FALSE`.

```{r}
#| label: grass
#| echo: false
#| fig-alt:
#|   - "Raster map of volcanic elevation using the GRASS aspect palette. Fill color encodes elevation across the volcanic terrain."
#|   - "Raster map of volcanic elevation using the GRASS aspectcolr palette. Fill color encodes elevation across the volcanic terrain."
#|   - "Raster map of volcanic elevation using the GRASS bcyr palette. Fill color encodes elevation across the volcanic terrain."
#|   - "Raster map of volcanic elevation using the GRASS bgyr palette. Fill color encodes elevation across the volcanic terrain."
#|   - "Raster map of volcanic elevation using the GRASS blues palette. Fill color encodes elevation across the volcanic terrain."
#|   - "Raster map of volcanic elevation using the GRASS byg palette. Fill color encodes elevation across the volcanic terrain."
#|   - "Raster map of volcanic elevation using the GRASS byr palette. Fill color encodes elevation across the volcanic terrain."
#|   - "Raster map of volcanic elevation using the GRASS celsius palette. Fill color encodes elevation across the volcanic terrain."
#|   - "Raster map of volcanic elevation using the GRASS corine palette. Fill color encodes elevation across the volcanic terrain."
#|   - "Raster map of volcanic elevation using the GRASS curvature palette. Fill color encodes elevation across the volcanic terrain."
#|   - "Raster map of volcanic elevation using the GRASS differences palette. Fill color encodes elevation across the volcanic terrain."
#|   - "Raster map of volcanic elevation using the GRASS elevation palette. Fill color encodes elevation across the volcanic terrain."
#|   - "Raster map of volcanic elevation using the GRASS etopo2 palette. Fill color encodes elevation across the volcanic terrain."
#|   - "Raster map of volcanic elevation using the GRASS evi palette. Fill color encodes elevation across the volcanic terrain."
#|   - "Raster map of volcanic elevation using the GRASS fahrenheit palette. Fill color encodes elevation across the volcanic terrain."
#|   - "Raster map of volcanic elevation using the GRASS forest_cover palette. Fill color encodes elevation across the volcanic terrain."
#|   - "Raster map of volcanic elevation using the GRASS gdd palette. Fill color encodes elevation across the volcanic terrain."
#|   - "Raster map of volcanic elevation using the GRASS grass palette. Fill color encodes elevation across the volcanic terrain."
#|   - "Raster map of volcanic elevation using the GRASS greens palette. Fill color encodes elevation across the volcanic terrain."
#|   - "Raster map of volcanic elevation using the GRASS grey palette. Fill color encodes elevation across the volcanic terrain."
#|   - "Raster map of volcanic elevation using the GRASS gyr palette. Fill color encodes elevation across the volcanic terrain."
#|   - "Raster map of volcanic elevation using the GRASS haxby palette. Fill color encodes elevation across the volcanic terrain."
#|   - "Raster map of volcanic elevation using the GRASS inferno palette. Fill color encodes elevation across the volcanic terrain."
#|   - "Raster map of volcanic elevation using the GRASS kelvin palette. Fill color encodes elevation across the volcanic terrain."
#|   - "Raster map of volcanic elevation using the GRASS magma palette. Fill color encodes elevation across the volcanic terrain."
#|   - "Raster map of volcanic elevation using the GRASS ndvi palette. Fill color encodes elevation across the volcanic terrain."
#|   - "Raster map of volcanic elevation using the GRASS ndwi palette. Fill color encodes elevation across the volcanic terrain."
#|   - "Raster map of volcanic elevation using the GRASS nlcd palette. Fill color encodes elevation across the volcanic terrain."
#|   - "Raster map of volcanic elevation using the GRASS oranges palette. Fill color encodes elevation across the volcanic terrain."
#|   - "Raster map of volcanic elevation using the GRASS plasma palette. Fill color encodes elevation across the volcanic terrain."
#|   - "Raster map of volcanic elevation using the GRASS population palette. Fill color encodes elevation across the volcanic terrain."
#|   - "Raster map of volcanic elevation using the GRASS population_dens palette. Fill color encodes elevation across the volcanic terrain."
#|   - "Raster map of volcanic elevation using the GRASS precipitation palette. Fill color encodes elevation across the volcanic terrain."
#|   - "Raster map of volcanic elevation using the GRASS precipitation_daily palette. Fill color encodes elevation across the volcanic terrain."
#|   - "Raster map of volcanic elevation using the GRASS precipitation_monthly palette. Fill color encodes elevation across the volcanic terrain."
#|   - "Raster map of volcanic elevation using the GRASS rainbow palette. Fill color encodes elevation across the volcanic terrain."
#|   - "Raster map of volcanic elevation using the GRASS ramp palette. Fill color encodes elevation across the volcanic terrain."
#|   - "Raster map of volcanic elevation using the GRASS reds palette. Fill color encodes elevation across the volcanic terrain."
#|   - "Raster map of volcanic elevation using the GRASS roygbiv palette. Fill color encodes elevation across the volcanic terrain."
#|   - "Raster map of volcanic elevation using the GRASS rstcurv palette. Fill color encodes elevation across the volcanic terrain."
#|   - "Raster map of volcanic elevation using the GRASS ryb palette. Fill color encodes elevation across the volcanic terrain."
#|   - "Raster map of volcanic elevation using the GRASS ryg palette. Fill color encodes elevation across the volcanic terrain."
#|   - "Raster map of volcanic elevation using the GRASS sepia palette. Fill color encodes elevation across the volcanic terrain."
#|   - "Raster map of volcanic elevation using the GRASS slope palette. Fill color encodes elevation across the volcanic terrain."
#|   - "Raster map of volcanic elevation using the GRASS soilmoisture palette. Fill color encodes elevation across the volcanic terrain."
#|   - "Raster map of volcanic elevation using the GRASS srtm palette. Fill color encodes elevation across the volcanic terrain."
#|   - "Raster map of volcanic elevation using the GRASS srtm_plus palette. Fill color encodes elevation across the volcanic terrain."
#|   - "Raster map of volcanic elevation using the GRASS terrain palette. Fill color encodes elevation across the volcanic terrain."
#|   - "Raster map of volcanic elevation using the GRASS viridis palette. Fill color encodes elevation across the volcanic terrain."
#|   - "Raster map of volcanic elevation using the GRASS water palette. Fill color encodes elevation across the volcanic terrain."
#|   - "Raster map of volcanic elevation using the GRASS wave palette. Fill color encodes elevation across the volcanic terrain."
#| layout-ncol: 2
#| lightbox:
#|   group: r-grass
#|   description: GRASS color table palettes in tidyterra.
allpals <- sort(unique(grass_db$pal))

for (i in allpals) {
  ss <- baseplot +
    scale_fill_grass_c(palette = i, use_grass_range = FALSE) +
    labs(title = paste0(i)) +
    theme_void() +
    theme(plot.background = element_rect(fill = "white"))

  print(ss)
}
```

## `scale_fill_princess_*`

See `scale_fill_princess_c()` for details.

```{r}
#| label: princess
#| echo: false
#| fig-alt:
#|   - "Raster map of volcanic elevation using the Princess america palette. Fill color encodes elevation across the volcanic terrain."
#|   - "Raster map of volcanic elevation using the Princess arabia palette. Fill color encodes elevation across the volcanic terrain."
#|   - "Raster map of volcanic elevation using the Princess asia palette. Fill color encodes elevation across the volcanic terrain."
#|   - "Raster map of volcanic elevation using the Princess aura palette. Fill color encodes elevation across the volcanic terrain."
#|   - "Raster map of volcanic elevation using the Princess bell palette. Fill color encodes elevation across the volcanic terrain."
#|   - "Raster map of volcanic elevation using the Princess cold palette. Fill color encodes elevation across the volcanic terrain."
#|   - "Raster map of volcanic elevation using the Princess denmark palette. Fill color encodes elevation across the volcanic terrain."
#|   - "Raster map of volcanic elevation using the Princess ella palette. Fill color encodes elevation across the volcanic terrain."
#|   - "Raster map of volcanic elevation using the Princess france palette. Fill color encodes elevation across the volcanic terrain."
#|   - "Raster map of volcanic elevation using the Princess maori palette. Fill color encodes elevation across the volcanic terrain."
#|   - "Raster map of volcanic elevation using the Princess neworleans palette. Fill color encodes elevation across the volcanic terrain."
#|   - "Raster map of volcanic elevation using the Princess norge palette. Fill color encodes elevation across the volcanic terrain."
#|   - "Raster map of volcanic elevation using the Princess punz palette. Fill color encodes elevation across the volcanic terrain."
#|   - "Raster map of volcanic elevation using the Princess scotland palette. Fill color encodes elevation across the volcanic terrain."
#|   - "Raster map of volcanic elevation using the Princess snow palette. Fill color encodes elevation across the volcanic terrain."
#| layout-ncol: 2
#| lightbox:
#|   group: r-princess
#|   description: Princess palettes in tidyterra.
allpals <- sort(unique(princess_db$pal))

for (i in allpals) {
  ss <- baseplot +
    scale_fill_princess_c(palette = i) +
    labs(title = paste0(i)) +
    theme_void() +
    theme(plot.background = element_rect(fill = "white"))

  print(ss)
}
```
